Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-12-06 manuscript/supplemental_material.Rnw
fix acronyms
add missing \clearpage commands
use short captions for supplemental figures

        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        display.table[,`Human Name`:=NULL]
        # setnames(display.table,"Human Symbol","\\multicolumn{1}{p{0.5in}}{Human Symbol}")6
        setnames(display.table,"Fold Change","$\\text{log}_2$ FC")
        print(xtable(display.table,
                     display=c("s","s","e","g","f"),
                     caption=caption,
                     label=label,
                     align=c("c","p{0.7in}","r","r","r")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              tabular.environment=ifelse(is.na(members),"longtable","tabular"),
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              size="\\small",
              include.rownames=FALSE)

@@

<<human.lineage.specific,echo=FALSE,results="asis">>=
hum.all.lin.spec <-
    placenta.lineage.specific.table("hum.all",
                                    lineage.table=human.mouse.lineage.specific.analysis,
                                    members=NA,
                                    caption="\\lineagetableexplanationtypeiii{\\human}{all other species}",
                                    label="tab:human_lineage_specific_all"
                                    )
setkey(hum.all.lin.spec,"Human Symbol")
@

@@

combined.placenta.go.results <-
    placenta.go.results.permissive[,list(`GO ID`,`Permissive FDR`)
                                   ][placenta.go.results][order(Ontology,FDR),]
combined.placenta.go.results <-
    combined.placenta.go.results[,`p value`:=as.numeric(`p value`)]
setcolorder(combined.placenta.go.results,
            c("Ontology",
              "GO ID",
              "GO Term",
              "Core",
              "Expected",
              "p value",
              "FDR",
              "Permissive FDR"))
combined.placenta.go.results <-
    combined.placenta.go.results[,`p value`:=NULL]
setnames(combined.placenta.go.results,"Ontology","Ont.")
setnames(combined.placenta.go.results,"Expected","Exp.")
print(xtable(combined.placenta.go.results,
             digits=2,
             display=c("d","s","s","s","d","f","e","e")),
      include.rownames=FALSE,
      booktabs=TRUE,
      math.style.exponents=TRUE,
      floating=FALSE
      )
@ 
\caption{GO terms most significantly enriched
  (\Sexpr{placenta.go.results[,sum(Ontology=="MF")]} MF,
  \Sexpr{placenta.go.results[,sum(Ontology=="BP")]} BP) in
  the core placenta transcriptome ($\text{FPKM}>10$ in 14
  of 14 species) as compared to all genes with 1:1 human orthologs in
  all species. “Ont.” is the GO Ontology, where MF is molecular function 
  and BP is biological process; “Core” is the number of genes in the core
  transcriptome for each GO term; “Exp.” is the number of genes
  expected in the core transcriptome for each term; 
  “FDR” is the false discovery rate as estimated by the
  Benjamini and Hochberg procedure \cite{Benjamini1995:bh_procedure} based on
  p values calculated by the Fisher Exact Test; 
  “Permissive FDR” is the FDR resulting from the equivalent GO
  analysis using the permissive core placenta transcriptome 
  (1:1 orthologs with expression of at least 10 FPKM at least 10 of 14 species) 
  showing that the enriched GO terms are relatively insensitive to the precise
  set of species included.}
\label{tab:core_placenta_go_terms}
\end{table}

@@

\begin{table}
<<attachment_shape_significant,echo=FALSE,results="asis">>=
placenta_class_table("shape")
@ 
\caption{Top 50 genes which are associated with the shape of placenta attachment}
\label{tab:shape_genes}
\end{table}

@@

\begin{table}[h]
<<intimacy_significant,echo=FALSE,results="asis">>=
placenta_class_table("intimacy")
@ 
\caption{Top 50 genes which are associated with the degree of placenta intimacy}
\label{tab:intimacy_genes}
\end{table}

\begin{table}[h]
<<intimacy_significant_polr,echo=FALSE,results="asis">>=
placenta_class_table("intimacy",analysis.type="polr")
@ 
\caption{Top 50 genes which are associated with the degree of placenta intimacy; ordered analysis}
\label{tab:intimacy_genes_polr}
\end{table}

\subsection{Maternal-Fetal Interdigitation}

\begin{table}[h]
<<interdigitation_significant,echo=FALSE,results="asis">>=
placenta_class_table("interdigitation")
@ 
\caption{Top 50 genes which are associated with the degree of maternal-fetal interdigitation}
\label{tab:interdigitation_genes}
\end{table}

\begin{table}[h]
<<interdigitation_significant_polr,echo=FALSE,results="asis">>=
placenta_class_table("interdigitation",analysis.type="polr")
@ 
\caption{Top 50 genes which are associated with the degree of maternal-fetal interdigitation; ordered analysis}
\label{tab:interdigitation_genes_polr}
\end{table}

>>>>

        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        display.table[,`Human Name`:=NULL]
        # setnames(display.table,"Human Symbol","\\multicolumn{1}{p{0.5in}}{Human Symbol}")6
        setnames(display.table,"Fold Change","$\\text{log}_2$ FC")
        print(xtable(display.table,
                     display=c("s","s","e","g","f"),
                     caption=caption,
                     label=label,
                     align=c("c","p{0.7in}","r","r","r")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              caption.placement="top",
              tabular.environment=ifelse(is.na(members),"longtable","tabular"),
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              size="\\small",
              include.rownames=FALSE)

@@

<<human.lineage.specific,echo=FALSE,results="asis">>=
hum.all.lin.spec <-
    placenta.lineage.specific.table("hum.all",
                                    lineage.table=human.mouse.lineage.specific.analysis,
                                    members=NA,
                                    caption=c("\\lineagetableexplanationtypeiii{\\human}{all other species}","\\human lineage specific genes")
                                    label="tab:human_lineage_specific_all"
                                    )
setkey(hum.all.lin.spec,"Human Symbol")
@

@@

combined.placenta.go.results <-
    placenta.go.results.permissive[,list(`GO ID`,`Permissive FDR`)
                                   ][placenta.go.results][order(Ontology,FDR),]
combined.placenta.go.results <-
    combined.placenta.go.results[,`p value`:=as.numeric(`p value`)]
setcolorder(combined.placenta.go.results,
            c("Ontology",
              "GO ID",
              "GO Term",
              "Core",
              "Expected",
              "p value",
              "FDR",
              "Permissive FDR"))
combined.placenta.go.results <-
    combined.placenta.go.results[,`p value`:=NULL]
setnames(combined.placenta.go.results,"Ontology","Ont.")
setnames(combined.placenta.go.results,"Expected","Exp.")
print(xtable(combined.placenta.go.results,
             digits=2,
             display=c("d","s","s","s","d","f","e","e")),
      include.rownames=FALSE,
      booktabs=TRUE,
      math.style.exponents=TRUE,
      floating=FALSE
      )
@ 
\caption[Core placenta GO Enrichment]{GO terms most significantly enriched
  (\Sexpr{placenta.go.results[,sum(Ontology=="MF")]} MF,
  \Sexpr{placenta.go.results[,sum(Ontology=="BP")]} BP) in
  the core placenta transcriptome ($\text{FPKM}>10$ in 14
  of 14 species) as compared to all genes with 1:1 human orthologs in
  all species. “Ont.” is the GO Ontology, where MF is molecular function 
  and BP is biological process; “Core” is the number of genes in the core
  transcriptome for each GO term; “Exp.” is the number of genes
  expected in the core transcriptome for each term; 
  “FDR” is the false discovery rate as estimated by the
  Benjamini and Hochberg procedure \cite{Benjamini1995:bh_procedure} based on
  p values calculated by the Fisher Exact Test; 
  “Permissive FDR” is the FDR resulting from the equivalent GO
  analysis using the permissive core placenta transcriptome 
  (1:1 orthologs with expression of at least 10 FPKM at least 10 of 14 species) 
  showing that the enriched GO terms are relatively insensitive to the precise
  set of species included.}
\label{tab:core_placenta_go_terms}
\end{table}

@@

\begin{table}[h]
<<attachment_shape_significant,echo=FALSE,results="asis">>=
placenta_class_table("shape")
@ 
\caption[Genes associated with attachment]{Top 50 genes which are associated with the shape of placenta attachment}
\label{tab:shape_genes}
\end{table}

@@

\begin{table}[h]
<<intimacy_significant,echo=FALSE,results="asis">>=
placenta_class_table("intimacy")
@ 
\caption[Genes associated with intimacy]{Top 50 genes which are associated with the degree of placenta intimacy}
\label{tab:intimacy_genes}
\end{table}

\begin{table}[h]
<<intimacy_significant_polr,echo=FALSE,results="asis">>=
placenta_class_table("intimacy",analysis.type="polr")
@ 
\caption[Intimacy genes: ordered analysis]{Top 50 genes which are associated with the degree of placenta intimacy; ordered analysis}
\label{tab:intimacy_genes_polr}
\end{table}

\clearpage
\subsection{Maternal-Fetal Interdigitation}

\begin{table}[h]
<<interdigitation_significant,echo=FALSE,results="asis">>=
placenta_class_table("interdigitation")
@ 
\caption[Genes associated with interdigitation]{Top 50 genes which are associated with the degree of maternal-fetal interdigitation}
\label{tab:interdigitation_genes}
\end{table}

\begin{table}[h]
<<interdigitation_significant_polr,echo=FALSE,results="asis">>=
placenta_class_table("interdigitation",analysis.type="polr")
@ 
\caption[Interdigitation genes: ordered analysis]{Top 50 genes which are associated with the degree of maternal-fetal interdigitation; ordered analysis}
\label{tab:interdigitation_genes_polr}
\end{table}
